title: "Crash Data: City of Tempe"
output: 
  flexdashboard::flex_dashboard:
    theme: yeti
    source: embed
    smart: false
    runtime: shiny
    orientation: rows
    
    
---

```{r global, include=FALSE}
library( flexdashboard )
library( tidyverse )
library( ggmap )
library( leaflet )
library( viridis )
library( shiny )
library( DT )
library( pander )
library( knitr )
library( rsconnect )


URL <- "https://github.com/DS4PS/Data-Science-Class/blob/master/DATA/TempeTrafficAccidents.rds?raw=true"
dat <- readRDS(gzcon(url( URL )))

dat <- na.omit(dat) # omit any rows with NAs
dat$fatal <- dat$Totalfatalities > 0 
dat$inj <- dat$Totalinjuries > 0 & dat$Totalfatalities == 0
dat$nohurt <- dat$Totalfatalities + dat$Totalinjuries == 0

date.vec   <- strptime( dat$DateTime, format="%m/%d/%y %H:%M" )
dat$hour   <- format( date.vec, format="%H" ) %>% as.numeric()
dat$month  <- format( date.vec, format="%b" )
dat$day    <- format( date.vec, format="%a" )
dat$day365 <- format( date.vec, format="%j" )
dat$week   <- format( date.vec, format="%V" )

dat <- 
  dat %>% 
  mutate( time.of.day = case_when( hour >= 6 & hour <= 9 ~ "Morning Commute", 
                                   hour >= 16 & hour <= 19 ~ "Evening Commute", 
                                   hour >= 14 & hour <= 15 ~ "School Pickup", 
                                   hour >= 9 & hour <= 13 ~ "Work", 
                                   hour >= 20 & hour <= 23 ~ "Night", 
                                   hour <= 5 & hour >= 0 ~ "Midnight to Dawn") )

dat$harm <- ifelse( dat$Totalinjuries > 0 | dat$Totalfatalities > 0, "Harm", "No Harm" )

dat <- 
  dat %>% 
  mutate( d1.substance = case_when( AlcoholUse_Drv1 == "Alcohol" & 
                                      DrugUse_Drv1 == "No Apparent Influence" ~ "Alcohol", 
                                   AlcoholUse_Drv1 == "No Apparent Influence" & 
                                     DrugUse_Drv1 == "Drugs" ~ "Drugs", 
                                   AlcoholUse_Drv1 == "Alcohol" & 
                                     DrugUse_Drv1 == "Drugs" ~ "Alcohol and Drugs", 
                                   AlcoholUse_Drv1 == "No Apparent Influence" & 
                                     DrugUse_Drv1 == "No Apparent Influence" ~ "No Apparent Influence"))

dat <- 
  dat %>% 
  mutate( d2.substance = case_when( AlcoholUse_Drv2 == "Alcohol" & 
                                      DrugUse_Drv2 == "No Apparent Influence" ~ "Alcohol", 
                                    AlcoholUse_Drv2 == "No Apparent Influence" & 
                                      DrugUse_Drv2 == "Drugs" ~ "Drugs", 
                                    AlcoholUse_Drv2 == "Alcohol" & 
                                      DrugUse_Drv2 == "Drugs" ~ "Alcohol and Drugs", 
                                    AlcoholUse_Drv2 == "No Apparent Influence" & 
                                      DrugUse_Drv2 == "No Apparent Influence" ~ "No Apparent Influence"))

dat$age.cat1 <- case_when( dat$Age_Drv1 >= 0 & 
                            dat$Age_Drv1 <= 18 ~ "Youth", 
                          dat$Age_Drv1 >= 19 & 
                            dat$Age_Drv1 <= 25 ~ "Young Adult", 
                          dat$Age_Drv1 >= 26 & 
                            dat$Age_Drv1 <= 64 ~ "Adult", 
                          dat$Age_Drv1 >= 65 ~ "Senior")

dat$age.cat2 <- case_when( dat$Age_Drv2 >= 0 & 
                            dat$Age_Drv2 <= 18 ~ "Youth", 
                          dat$Age_Drv2 >= 19 & 
                            dat$Age_Drv2 <= 25 ~ "Young Adult", 
                          dat$Age_Drv2 >= 26 & 
                            dat$Age_Drv2 <= 64 ~ "Adult", 
                          dat$Age_Drv2 >= 65 ~ "Senior")




```



Traffic Accidents By Day and Time {data-navmenu="Analysis"}
=====================================  


Inputs {.sidebar}
-------------------------------------

```{r}

checkboxGroupInput("days", label = h3("Day of Week"), 
    choices = list("Monday"    = "Mon", 
                   "Tuesday"   = "Tue", 
                   "Wednesday" = "Wed", 
                   "Thursday"  = "Thu",
                   "Friday"    = "Fri",
                   "Saturday"  = "Sat",
                   "Sunday"    = "Sun" ),
    selected = c("Fri","Sat","Sun"))

sliderInput("hour", 
            label = h3("Time of Day"),
            min = 0, 
            max = 23,
            value = c(6, 12))

# parameters

```

   
Outputs
-------------------------------------

### Traffic Accidents By Day & Time


```{r}

#leaflet
renderLeaflet({
  
  days.of.week <- input$days    # vector will all checked values
  start.time <- input$hour[1]   # sliderInput lower value
  end.time  <-  input$hour[2] 
  
  d2 <-
    dat %>%
    filter( day %in% input$days, 
            hour >= start.time & hour <= end.time )
  
  d2$col.vec <- ifelse( d2$nohurt, "gray20", ifelse(d2$inj, "steelblue", "darkorange") )              
    
  point.size <- d2$Totalinjuries + d2$Totalfatalities

  crash.details <- paste0( "Time: ", d2$DateTime, "",
                     "Total Fatalities: ", d2$Totalfatalities, "",
                     "Total Injuries: ", d2$Totalinjuries, "",
                     "Collision type: ", d2$Collisionmanner)
  
  tempe <- leaflet( ) %>% 
              addProviderTiles( "CartoDB.Positron" )  %>%
              setView( lng=-111.9278, lat=33.39951, zoom=13 )
  
  
  addCircles( tempe, lng=d2$Longitude, lat=d2$Latitude,
              fillColor=d2$col.vec, fillOpacity=0.5, 
              stroke=F, radius=50*(1+0.33*point.size),
              popup=crash.details )


})
```   



Driver Characteristics {data-navmenu="Analysis"}
=====================================  

Sidebar {.sidebar}
-------------------------------------
Driver Characteristics
```{r}
sliderInput("d1age", label = h4("Driver 1 Age"), 
            min = 15, max = 100, value = c(18,36) )
sliderInput("d2age", label = h4("Driver 2 Age"), 
            min = 15, max = 100, value = c(18,36) )
selectInput("d1gender", label = h4("Driver 1 Gender"), 
    choices = c("Male","Female", "Unknown"), selected = c("Male"))
selectInput("d2gender", label = h4("Driver 2 Gender"), 
    choices = c("Male","Female", "Unknown"), selected = c("Male"))
radioButtons("d1pedcy", label = h4("Driver 1 Transportation"),
    choices = c("Driver", "Pedalcyclist", "Pedestrian"), selected = c("Driver"))
radioButtons("d2pedcy", label = h4("Driver 2 Transportation"),
    choices = c("Driver", "Pedalcyclist", "Pedestrian"), selected = c("Driver"))

```





row
-------------------------------------


### Number of Crashes
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter( Age_Drv1 >= input$d1age[1] & Age_Drv1 <= input$d1age[2], 
            Age_Drv2 >= input$d2age[1] & Age_Drv2 <= input$d2age[2], 
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender, 
            Unittype_One %in% input$d1pedcy, 
            Unittype_Two %in% input$d2pedcy )
  
  crashes <- count( d2 )
  valueBox(crashes, 
           icon = "fa-pencil",
           color = ifelse( crashes > 50, "danger", " alright") )
})
```


### Total Injuries
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter( Age_Drv1 >= input$d1age[1] & Age_Drv1 <= input$d1age[2], 
            Age_Drv2 >= input$d2age[1] & Age_Drv2 <= input$d2age[2], 
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender, 
            Unittype_One %in% input$d1pedcy, 
            Unittype_Two %in% input$d2pedcy )
  
  total.injuries <- sum( d2$Totalinjuries )
  valueBox(total.injuries, 
           icon = "fa-angry",
           color = ifelse( total.injuries > 30, "danger", " alright" ))
})
```

### Total Fatalities
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter( Age_Drv1 >= input$d1age[1] & Age_Drv1 <= input$d1age[2], 
            Age_Drv2 >= input$d2age[1] & Age_Drv2 <= input$d2age[2], 
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender, 
            Unittype_One %in% input$d1pedcy, 
            Unittype_Two %in% input$d2pedcy )

  total.fatalities <- sum( d2$Totalfatalities )
  valueBox( total.fatalities, 
            icon = "fa-briefcase-medical",
            color = ifelse(total.fatalities > 10, "danger", " alright"))
})
```


### Rate of Harm
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter( Age_Drv1 >= input$d1age[1] & Age_Drv1 <= input$d1age[2], 
            Age_Drv2 >= input$d2age[1] & Age_Drv2 <= input$d2age[2], 
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender, 
            Unittype_One %in% input$d1pedcy, 
            Unittype_Two %in% input$d2pedcy )
  
  rate.of.harm <- round(length(which(d2$harm == "Harm"))/count(d2), 3)
  valueBox(rate.of.harm, 
           icon = "fa-pencil",
           color = ifelse(rate.of.harm > 0.5, "danger", " alright"))
})
```




   
Outputs
-------------------------------------

### Traffic Accidents by Driver Characteristics


```{r}

renderLeaflet({
 
  
  d2 <-
    dat %>%
    filter( Age_Drv1 >= input$d1age[1] & Age_Drv1 <= input$d1age[2], 
            Age_Drv2 >= input$d2age[1] & Age_Drv2 <= input$d2age[2], 
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender, 
            Unittype_One %in% input$d1pedcy, 
            Unittype_Two %in% input$d2pedcy )
  
  d2$col.vec <- ifelse( d2$nohurt, "gray20", ifelse(d2$inj, "steelblue", "darkorange") )              
    
  point.size <- d2$Totalinjuries + d2$Totalfatalities

  crash.details <- paste0( "Time: ", d2$DateTime, "",
                     "Total Fatalities: ", d2$Totalfatalities, "",
                     "Total Injuries: ", d2$Totalinjuries, "",
                     "Collision type: ", d2$Collisionmanner)
  
  tempe <- leaflet( ) %>% 
              addProviderTiles( "CartoDB.Positron" )  %>%
              setView( lng=-111.9278, lat=33.39951, zoom=13 )
  
  
  addCircles( tempe, lng=d2$Longitude, lat=d2$Latitude,
              fillColor=d2$col.vec, fillOpacity=0.5, 
              stroke=F, radius=50*(1+0.33*point.size),
              popup=crash.details )


})
```   






Drivers 2 {data-navmenu="Analysis"}
=====================================  

Sidebar {.sidebar}
-------------------------------------
Driver Characteristics
```{r}
sliderInput("driver.1.age", label = h4("Driver 1 Age"), 
            min = 15, max = 100, value = c(18,36) )
sliderInput("driver.2.age", label = h4("Driver 2 Age"), 
            min = 15, max = 100, value = c(18,36) )
selectInput("driver.1.gender", label = h4("Driver 1 Gender"), 
    choices = c("Male","Female", "Unknown"), selected = c("Male"))
selectInput("driver.2.gender", label = h4("Driver 2 Gender"), 
    choices = c("Male","Female", "Unknown"), selected = c("Male"))
radioButtons("driver.1.pedcy", label = h4("Driver 1 Transportation"),
    choices = c("Driver", "Pedalcyclist", "Pedestrian"), selected = c("Driver"))
radioButtons("driver.2.pedcy", label = h4("Driver 2 Transportation"),
    choices = c("Driver", "Pedalcyclist", "Pedestrian"), selected = c("Driver"))

```




-------------------------------------


### Number of Crashes
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter( Age_Drv1 >= input$d1age[1] & Age_Drv1 <= input$d1age[2], 
            Age_Drv2 >= input$d2age[1] & Age_Drv2 <= input$d2age[2], 
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender, 
            Unittype_One %in% input$d1pedcy, 
            Unittype_Two %in% input$d2pedcy )
  
  crashes <- count( d2 )
  valueBox(crashes, 
           icon = "fa-pencil",
           color = ifelse( crashes > 50, "danger", " alright") )
})
```


### Total Injuries
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter( Age_Drv1 >= input$d1age[1] & Age_Drv1 <= input$d1age[2], 
            Age_Drv2 >= input$d2age[1] & Age_Drv2 <= input$d2age[2], 
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender, 
            Unittype_One %in% input$d1pedcy, 
            Unittype_Two %in% input$d2pedcy )
  
  total.injuries <- sum( d2$Totalinjuries )
  valueBox(total.injuries, 
           icon = "fa-angry",
           color = ifelse( total.injuries > 30, "danger", " alright" ))
})
```

### Total Fatalities
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter( Age_Drv1 >= input$d1age[1] & Age_Drv1 <= input$d1age[2], 
            Age_Drv2 >= input$d2age[1] & Age_Drv2 <= input$d2age[2], 
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender, 
            Unittype_One %in% input$d1pedcy, 
            Unittype_Two %in% input$d2pedcy )

  total.fatalities <- sum( d2$Totalfatalities )
  valueBox( total.fatalities, 
            icon = "fa-briefcase-medical",
            color = ifelse(total.fatalities > 10, "danger", " alright"))
})
```


### Rate of Harm
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter( Age_Drv1 >= input$d1age[1] & Age_Drv1 <= input$d1age[2], 
            Age_Drv2 >= input$d2age[1] & Age_Drv2 <= input$d2age[2], 
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender, 
            Unittype_One %in% input$d1pedcy, 
            Unittype_Two %in% input$d2pedcy )
  
  rate.of.harm <- round(length(which(d2$harm == "Harm"))/count(d2), 3)
  valueBox(rate.of.harm, 
           icon = "fa-pencil",
           color = ifelse(rate.of.harm > 0.5, "danger", " alright"))
})
```




   
Column
-------------------------------------

### Driver 1


```{r}

renderLeaflet({
  
  # days.of.week <- input$days    # vector will all checked values
  # start.time <- input$hour[1]   # sliderInput lower value
  # end.time  <-  input$hour[2] 
  
  d10 <-
    dat %>%
    filter( Age_Drv1 >= input$driver.1.age[1] & Age_Drv1 <= input$driver.1.age[2], 
            Gender_Drv1 %in% input$driver.1.gender, 
            Unittype_One %in% input$driver.1.pedcy )
  
  d10$col.vec <- ifelse( d10$nohurt, "gray20", ifelse(d10$inj, "steelblue", "darkorange") )              
    
  point.size <- d10$Totalinjuries + d10$Totalfatalities

  crash.details <- paste0( "Time: ", d10$DateTime, "",
                     "Total Fatalities: ", d10$Totalfatalities, "",
                     "Total Injuries: ", d10$Totalinjuries, "",
                     "Collision type: ", d10$Collisionmanner)
  
  tempe <- leaflet( ) %>% 
              addProviderTiles( "CartoDB.Positron" )  %>%
              setView( lng=-111.9278, lat=33.39951, zoom=13 )
  
  
  addCircles( tempe, lng=d10$Longitude, lat=d10$Latitude,
              fillColor=d10$col.vec, fillOpacity=0.5, 
              stroke=F, radius=50*(1+0.33*point.size),
              popup=crash.details )


})
```   



### Driver 2


```{r}

renderLeaflet({
  
  # days.of.week <- input$days    # vector will all checked values
  # start.time <- input$hour[1]   # sliderInput lower value
  # end.time  <-  input$hour[2] 
  
  d11 <-
    dat %>%
    filter( Age_Drv2 >= input$driver.2.age[1] & Age_Drv2 <= input$driver.2.age[2], 
            Gender_Drv2 %in% input$driver.2.gender, 
            Unittype_Two %in% input$driver.2.pedcy )
  
  d11$col.vec <- ifelse( d11$nohurt, "gray20", ifelse(d11$inj, "steelblue", "darkorange") )              
    
  point.size2 <- d11$Totalinjuries + d11$Totalfatalities

  crash.details2 <- paste0( "Time: ", d11$DateTime, "",
                     "Total Fatalities: ", d11$Totalfatalities, "",
                     "Total Injuries: ", d11$Totalinjuries, "",
                     "Collision type: ", d11$Collisionmanner)
  
  tempe2 <- leaflet( ) %>% 
              addProviderTiles( "CartoDB.Positron" )  %>%
              setView( lng=-111.9278, lat=33.39951, zoom=13 )
  
  
  addCircles( tempe2, lng=d11$Longitude, lat=d11$Latitude,
              fillColor=d11$col.vec, fillOpacity=0.5, 
              stroke=F, radius=50*(1+0.33*point.size2),
              popup=crash.details2 )


})
```   

 





















Accidents by Weather, Day, & Time {data-navmenu="Analysis"}
=====================================  


Inputs {.sidebar}
-------------------------------------

```{r}




shiny:: selectInput(inputId = "WDAA", 
                   label = h2("Weather During Accident"), 
    choices = c( "Clear", "Cloudy", "Unknown"),
    selected = ("Clear"))
 

shiny:: selectInput(inputId = "SCC", 
                   label = h2("Surface Condition"), 
    choices = c( "Dry", "Wet", "Unknown"),
    selected = ("Dry"))

selectInput("TOD", label = h2("Time of Day"), 
    choices = c("Morning Commute","Evening Commute", "School Pickup", "Work", "Night", "Midnight to Dawn"), selected = c("Work"))




# parameters

```

Although it would be reasonable to assume that there would be more accidents in wet or cloudy conditions, the opposite is true in Tempe. This is because the area has little [rainfall](https://www.maricopa.gov/625/Rainfall-Data) throughout the year.





   
Outputs
-------------------------------------

### Weather


```{r}

#leaflet
renderLeaflet({
  
  
  WDA <- input$WDAA  
  SC <- input$SCC
  start.time <- input$hour[1]   
  end.time  <-  input$hour[2] 
 
  
  
  d2 <-
    dat %>%
    filter( Weather ==  WDA , 
            SurfaceCondition == SC,
            hour >= start.time & hour <= end.time,
            time.of.day %in% input$TOD)
    

  point.size <- d2$Totalinjuries + d2$Totalfatalities

  crash.details <- paste0( "Time: ", d2$DateTime, "",
                     "Total Fatalities: ", d2$Totalfatalities, "",
                     "Total Injuries: ", d2$Totalinjuries, "",
                     "Collision type: ", d2$Collisionmanner,"",
                       "Travel Direction 1: ", d2$Traveldirection_One, "",
                    "Travel Direction 2: ", d2$Traveldirection_Two)
  
  

  tempe <- leaflet() %>% addTiles() %>% setView(lng=-111.9278, lat=33.39951, zoom=5) %>%
  addWMSTiles(
    "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913",
    options = WMSTileOptions(format = "image/png", transparent = TRUE),
    attribution = "Weather data © 2012 IEM Nexrad")
  
  addMarkers( tempe, lng=d2$Longitude, lat=d2$Latitude,
            
              popup=crash.details,clusterOptions =markerClusterOptions() )
  
  
  
  
  
})
```   





























Accidents by Age Category, Gender, & Year {data-navmenu="Analysis"}
=====================================  

Sidebar {.sidebar}
-------------------------------------

```{r}

checkboxGroupInput("d1agecat", label = h5("Driver 1 Age Category"), 
    choices = list("Youth", 
                   "Young Adult", 
                   "Adult", 
                  "Senior"),
    selected = c("Youth","Young Adult"))

checkboxGroupInput("d2agecat", label = h5("Driver 2 Age Category"), 
    choices = list("Youth", 
                   "Young Adult", 
                   "Adult", 
                  "Senior"),
    selected = c("Youth","Young Adult"))


sliderInput("yr", 
            label = h3("Year"),
            min = 2012, 
            max = 2018,
            value = c(2012, 2014))



```






Row 
-------------------------------------


### Number of Crashes
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter( 
            age.cat1 %in% input$d1agecat, 
            age.cat2 %in% input$d2agecat,
            Year %in% input$yr)
  
  crashes <- count( d2 )
  valueBox(crashes, 
           icon = "fa-sad-tear",
           color =  ifelse( crashes > 50, "danger", " alright") )
})
```


### Total Injuries
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter( 
            age.cat1 %in% input$d1agecat, 
            age.cat2 %in% input$d2agecat,
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender)
  
  total.injuries <- sum( d2$Totalinjuries )
  valueBox(total.injuries, 
           icon = "fa-hospital",
           color =  ifelse( total.injuries > 30, "danger", " alright" ))
})
```

### Total Fatalities
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter( 
            age.cat1 %in% input$d1agecat, 
            age.cat2 %in% input$d2agecat,
             Year %in% input$yr)

  total.fatalities <- sum( d2$Totalfatalities )
  valueBox( total.fatalities, 
            icon = "fa-skull",
            color =  ifelse(total.fatalities > 10, "danger", " alright"))
})
```


### Rate of Harm
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter( 
            age.cat1 %in% input$d1agecat, 
            age.cat2 %in% input$d2agecat,
             Year %in% input$yr)
  
  rate.of.harm <- round(length(which(d2$harm == "Harm"))/count(d2), 3)
  valueBox(rate.of.harm, 
           icon = "fa-procedures",
           color =  ifelse(rate.of.harm > 0.5, "danger", "alright"))
})





```

   
  
  
Outputs
-------------------------------------

### Traffic Accidents by Year & Age Category


```{r}

renderLeaflet({
  

  
  d2 <-
    dat %>%
    filter(  age.cat1 %in% input$d1agecat, 
            age.cat2 %in% input$d2agecat,
             Year %in% input$yr )
  
  d2$col.vec <- ifelse( d2$nohurt, "blue", ifelse(d2$inj, "green", "yellow") )              
    
  point.size <- d2$Totalinjuries + d2$Totalfatalities

  crash.details <- paste0( "Time: ", d2$DateTime, "",
                     "Total Fatalities: ", d2$Totalfatalities, "",
                     "Total Injuries: ", d2$Totalinjuries, "",
                     "Collision type: ", d2$Collisionmanner,"",
                       "Travel Direction 1: ", d2$Traveldirection_One, "",
                    "Travel Direction 2: ", d2$Traveldirection_Two)
  
  
  tempe <- leaflet() %>% addTiles() %>% setView(lng=-111.9278, lat=33.39951, zoom=13) %>%
  addWMSTiles(
    "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913",
    options = WMSTileOptions(format = "image/png", transparent = TRUE),
    attribution = "Weather data © 2012 IEM Nexrad")
  
  addMarkers( tempe, lng=d2$Longitude, lat=d2$Latitude,
            
              popup=crash.details,clusterOptions =markerClusterOptions() )

  
})
```   


















Accidents by Influence & Days of the Week {data-navmenu="Analysis"}
=====================================  

Sidebar {.sidebar}
-------------------------------------

```{r}

checkboxGroupInput("days2", label = h3("Day of Week"), 
    choices = list("Monday"    = "Mon", 
                   "Tuesday"   = "Tue", 
                   "Wednesday" = "Wed", 
                   "Thursday"  = "Thu",
                   "Friday"    = "Fri",
                   "Saturday"  = "Sat",
                   "Sunday"    = "Sun" ),
    selected = c("Fri","Sat","Sun"))


selectInput("d1influence", label = h4("Driver 1 Influence"), 
    choices = c("Alcohol","No Apparent Influence", "Drugs","Alcohol and Drugs"), selected = c("Alcohol"))

selectInput("d2influence", label = h4("Driver 2 Influence"), 
     choices = c("Alcohol","No Apparent Influence", "Drugs","Alcohol and Drugs"), selected = c("No Apparent Influence"))

```






Row 
-------------------------------------


### Number of Crashes
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter(
            day == input$days2,
            d1.substance %in% input$d1influence,
            d2.substance %in% input$d2influence
           )
  
  crashes <- count(d2)
  valueBox(crashes, 
           icon = "fa-sad-tear",
           color =  ifelse( crashes > 50, "danger", " alright") )
})
```


### Total Injuries
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter(  
            day %in% input$days2,
            d1.substance %in% input$d1influence,
            d2.substance %in% input$d2influence,
            AlcoholUse_Drv1 %in% input$d1influence,
            AlcoholUse_Drv2 %in% input$d2influence
             )
  
  total.injuries <- sum( d2$Totalinjuries )
  valueBox(total.injuries, 
           icon = "fa-hospital",
           color =  ifelse( total.injuries > 30, "danger", " alright" ))
})
```

### Total Fatalities
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter(  
            day %in% input$days2,
            d1.substance %in% input$d1influence,
            d2.substance %in% input$d2influence
             )

  total.fatalities <- sum( d2$Totalfatalities )
  valueBox( total.fatalities, 
            icon = "fa-skull",
            color =   ifelse(total.fatalities > 10, "danger", " alright"))
})
```


### Rate of Harm
  
```{r}
renderValueBox({
  d2 <-
    dat %>%
    filter(  
            day %in% input$days2,
            d1.substance %in% input$d1influence,
            d2.substance %in% input$d2influence
             )
  
  rate.of.harm <- round(length(which(d2$harm == "Harm"))/count(d2), 3)
  valueBox(rate.of.harm, 
           icon = "fa-procedures",
           color =  ifelse(rate.of.harm > 0.5, "danger", " alright"))
})
```




   
Column {.tabset .tabset-fade}
-------------------------------------

### Driver 1


```{r}

renderLeaflet({
  
 
  
  d10 <-
    dat %>%
    filter( 
            day %in% input$days2,
            d1.substance %in% input$d1influence,
              )
  
  d10$col.vec <- ifelse( d10$nohurt, "gray20", ifelse(d10$inj, "steelblue", "darkorange") )              
    
  point.size <- d10$Totalinjuries + d10$Totalfatalities

  crash.details12 <- paste0( "Time: ", d10$DateTime, "",
                     "Total Fatalities: ", d10$Totalfatalities, "",
                     "Total Injuries: ", d10$Totalinjuries, "",
                     "Collision type: ", d10$Collisionmanner,"",
                       "Travel Direction 1: ", d10$Traveldirection_One, "",
                    "Travel Direction 2: ", d10$Traveldirection_Two)
  
 
  tempe <- leaflet() %>% addTiles() %>% setView(lng=-111.9278, lat=33.39951, zoom=12) %>%
  addWMSTiles(
    "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913",
    options = WMSTileOptions(format = "image/png", transparent = TRUE),
    attribution = "Weather data © 2012 IEM Nexrad")
  
  addMarkers( tempe, lng=d10$Longitude, lat=d10$Latitude,
            
              popup=crash.details12,clusterOptions =markerClusterOptions() )


})
```   



### Driver 2


-------------------
```{r}

renderLeaflet({
  
  
  
  d11 <-
    dat %>%
    filter(  
            day %in% input$days2,
            d2.substance %in% input$d2influence )
  
  d11$col.vec <- ifelse( d11$nohurt, "gray20", ifelse(d11$inj, "steelblue", "darkorange") )              
    
  point.size2 <- d11$Totalinjuries + d11$Totalfatalities

  crash.details2 <- paste0( "Time: ", d11$DateTime, "",
                     "Total Fatalities: ", d11$Totalfatalities, "",
                     "Total Injuries: ", d11$Totalinjuries, "",
                     "Collision type: ", d11$Collisionmanner,"",
                       "Travel Direction 1: ", d11$Traveldirection_One, "",
                    "Travel Direction 2: ", d11$Traveldirection_Two)
  

  tempe <- leaflet() %>% addTiles() %>% setView(lng=-111.9278, lat=33.39951, zoom=12) %>%
  addWMSTiles(
    "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913",
    options = WMSTileOptions(format = "image/png", transparent = TRUE),
    attribution = "Weather data © 2012 IEM Nexrad")
  
  addMarkers( tempe, lng=d11$Longitude, lat=d11$Latitude,
            
              popup=crash.details2,clusterOptions =markerClusterOptions() )


})
```   







About
===================================== 



Row 
-------------------------------------

### About this Dashboard
This dashboard was created using the [City of Tempi Open Data Portal.] (https://data.tempe.gov/dataset/high-severity-traffic-crashes-1-08)






> The world is now awash in data and we can see consumers in a lot clearer ways. ~ Max Levchin, PayPal co-founder

### Dashboard Author
**Marina Chaji**, holds a Masters of Arts degree in Economics and has a background in research analytics. She enjoys learning more ways to help make the data she collects come alive. 


[Add me on Linkedin!](https://www.linkedin.com/in/marinachaji/)



Row
-------------------------------------

DATA DICTIONARY


```{r}
URL.dd <- "https://raw.githubusercontent.com/DS4PS/cpp-526-fall-2019/master/labs/final-project/TempeTrafficAccidentsDataDictionary.csv"
data.dictionary <- read.csv( URL.dd, stringsAsFactors=F )

data.dictionary$description <- stringi::stri_trans_general( data.dictionary$description, "latin-ascii" )
data.dictionary %>%
  select( column, description ) %>%
  pander( )
```



Data
=====================================  

```{r}
# library( DT )

these.buttons <- c( 'copy', 'csv', 'pdf', 'print' )

renderDataTable({
  datatable(dat[1:100,], filter='bottom', rownames=FALSE, 
           #options=list( pageLength=5, autoWidth=TRUE ),
           fillContainer=TRUE, 
           style="bootstrap",
           class='table-condensed table-striped',
           extensions = 'Buttons', 
           options=list( dom='Bfrtip', 
                         buttons=these.buttons  ) )
})
```



